W(ha/il)t's the (wo/ba)rd?

### 2022-04-30 paper_5.rmd
Fix typos and cross-ref issues

But, the above pair may also more succintly be represented in-text by a form like (PL/GR)\_\_\_\_ when more appropriate.

@@

## Data
### Corpora: What's a "word"?
One of the most common side effects of continued study in linguistics is a diminished ability to describe what defines a "word". For the purpose of this study, we have taken three lists that purport to contain a wide variety of English words, though certainly not their totality, and used them to generate all possible combinations that produce a valid Split Decisions word pair (in terms of a two-character split pair, etc.). As a result, many entries that appear in one or more of these corpora may seem esoteric, or even offensive, to be cosnidered as words; however, the only predicate for their inclusion in these lists is attested usage, not whether they prescriptively *ought* to be a word. Words appear in these corpora in lexicographical order (i.e., sorted alphabetically), one on each line, in lower-case letters.

@@

Before examining the splits and other puzzle-specific properties of words, we examined the distribution of one- and two-letter sequences, herefore *1-grams* and *2-grams*, respectively, to see if any Zipf-like patterns emerged. Surprisingly, however, the data outperformed Zipf's Law (i.e., contained many more observations than the power law predicts) in these categories. Figure \ref{fig:single-letter-distribution} below, which examines the total occurrences of each letter in all valid words, exemplifies this. The black dashed curve represents the expected count of each letter according to Zipf's Law, whereas the total height of each bar represents the actual count. As an example, the second-most common letter, *s*, is expected to occur $539,056 / 2 = 269,528$ times, or 50% of the total occurrences of *e*. In the data, it instead appears roughly 84% as often, a total of `r letter_rankings$total[[2]] |> format(big.mark = ",")` times. This disparity between Zipfian expectation and reality only increases in higher ranks, including the universally rare *q* and *j*, with a mean percent error of `r mean(letter_error$error) |> format(nsmall = 3)`% across all 26 letters.

@@

Again comparing the frequency distributions of splits, this time by position type instead of by corpus, against Zipf's Law yields another logarithmic correlation (note the log-scale in Figure \ref{fig:split-position-distribution}). That is, the observed distribution overperforms compared to the Zipf model before a particular inflection point, after which it begins to underperform with increasing error. Here, sheer combinatorics produces something of a visual confounding effect: middle splits outnumber both prefixes and suffixes by an order of magnitude, and encompass almost two hundred more 2-grams. As a result, we do not observe the same *x*-intercept as in Figure \ref{fig:two-letter-distribtion}, in which all three distributions seem to intersect their ideal Zipf curves (dashed lines) at the same rank. We again observe that prefixes and suffixes have this intersection at roughly the same point, around rank 255, but the middle-type splits have their inflection point around rank 445. By drawing dotted lines at the intersections of these distributions with the Zipf model, we see that the two intersections occur approximately 190 ranks apart and in both cases separate roughly 100 of the least frequent splits into a 'tail' of sorts.

>>>>

But, the above pair may also more succinctly be represented in-text by a form like (PL/GR)\_\_\_\_ when more appropriate.

@@

## Data
### Corpora: What's a "word"?
One of the most common side effects of continued study in linguistics is a diminished ability to describe what defines a "word". For the purpose of this study, we have taken three lists that purport to contain a wide variety of English words, though certainly not their totality, and used them to generate all possible combinations that produce a valid Split Decisions word pair (in terms of a two-character split pair, etc.). As a result, many entries that appear in one or more of these corpora may seem esoteric, or even offensive, to be considered as words; however, the only predicate for their inclusion in these lists is attested usage, not whether they prescriptively *ought* to be a word. Words appear in these corpora in lexicographical order (i.e., sorted alphabetically), one on each line, in lower-case letters.

@@

Before examining the splits and other puzzle-specific properties of words, we examined the distribution of one- and two-letter sequences, hereafter *1-grams* and *2-grams*, respectively, to see if any Zipf-like patterns emerged. Surprisingly, however, the data outperformed Zipf's Law (i.e., contained many more observations than the power law predicts) in these categories. Figure \ref{fig:single-letter-distribution} below, which examines the total occurrences of each letter in all valid words, exemplifies this. The black dashed curve represents the expected count of each letter according to Zipf's Law, whereas the total height of each bar represents the actual count. As an example, the second-most common letter, *s*, is expected to occur $539,056 / 2 = 269,528$ times, or 50% of the total occurrences of *e*. In the data, it instead appears roughly 84% as often, a total of `r letter_rankings$total[[2]] |> format(big.mark = ",")` times. This disparity between Zipfian expectation and reality only increases in higher ranks, including the universally rare *q* and *j*, with a mean percent error of `r mean(letter_error$error) |> format(nsmall = 3)`% across all 26 letters.

@@

Again comparing the frequency distributions of splits, this time by position type instead of by corpus, against Zipf's Law yields another logarithmic correlation (note the log-scale in Figure \ref{fig:split-position-distribution}). That is, the observed distribution overperforms compared to the Zipf model before a particular inflection point, after which it begins to underperform with increasing error. Here, sheer combinatorics produces something of a visual confounding effect: middle splits outnumber both prefixes and suffixes by an order of magnitude, and encompass almost two hundred more 2-grams. As a result, we do not observe the same *x*-intercept as in Figure \ref{fig:two-letter-distribution}, in which all three distributions seem to intersect their ideal Zipf curves (dashed lines) at the same rank. We again observe that prefixes and suffixes have this intersection at roughly the same point, around rank 255, but the middle-type splits have their inflection point around rank 445. By drawing dotted lines at the intersections of these distributions with the Zipf model, we see that the two intersections occur approximately 190 ranks apart and in both cases separate roughly 100 of the least frequent splits into a 'tail' of sorts.
